### 2023-07-14 python_excel_ite.tex
module positioning and structure section done

At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}). The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics. The PCDA course ends with intro to the \href{https://scikit-learn.org/stable/index.html}{scikit-learn library} for doing predictive modeling in Python and the AAP course begins where the PCDA course ends. The EwP module could also be used as part of a semester long Python based analytics course. Most AAP students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics. We do not strictly require the spreadsheet based course be taken before the PCDA or AAP courses as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python. The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab. Since 2020, these two courses have been also offered as online, asynchronous courses. The newer AAP course has been offered since 2021 in an online asynchronous mode. All three courses have \href{http://www.sba.oakland.edu/Faculty/isken/teaching.html}{extensive course websites} that are publicly accessible, including access to all of the video content and supporting files.

What kind of Python background do students need to have in order to be ready for the EwP module? A basic familiarity with Python fundamentals for data analysis is needed, including:

@@

The EwP module is made up of three submodules. The first of these focuses on building and using a typical spreadsheet model using Python. Then the functionality developed for doing this type of modeling in Python is deployed as a reusable package. We end with an introduction to using Python to manipulate Excel files.

%TODO needs much editing and shaping

All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks. Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{Jupyter_proposal} and allow one to create educational narratives that support iterative exploration and development. The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community \cite{perkel2018jupyter}. Jupyter notebooks can also include equations written in LaTex, data visualizations, and other multi-media. The technology is free, open source, runs on all platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.

Central to Jupyter notebooks is a mix of \textit{code cells} and \textit{markdown cells}. Code cells contain executable code while markdown cells contain plain text using \href{URL}{markdown} for styling text. For example, Figure \ref{cells_edit_model} shows two markdown cells separated by a code cell as they look when in editing mode. After running the cells, they appear as show in Figure \ref{cells_after_running}

As will be described below, Jupyter notebooks

Jupyter as an authoring environment - blog posts were written as Jupyter notebooks.

%TODO - more on structure of my Jupyter notebooks for teaching

%TODO - more on Jupyter project

>>>>

At our institution, The EwP module is part of a business course entitled ``Advanced Analytics with Python'' (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis6900}{AAP}). The students taking this course have already taken the PCDA course within which they learned fundamental Python programming within the context of data analytics over seven weeks. The PCDA course ends with intro to the \href{https://scikit-learn.org/stable/index.html}{scikit-learn library} for doing predictive modeling in Python and the AAP course begins where the PCDA course ends. The EwP module could also be used as part of a semester long Python based analytics course. Most AAP students have also already taken "Business Analytics" (\href{http://www.sba.oakland.edu/faculty/isken/courses/mis5460}{BA}) which is a spreadsheet based introduction to business analytics. We do not strictly require the spreadsheet based course be taken before the PCDA or AAP courses as that course draws students from outside of our business programs who are strictly interested in programmatic tools such as Python. The BA and PCDA courses have historically been offered in face to face mode in a computer teaching lab. Since 2020, these two courses have been also offered as online, asynchronous courses. The newer AAP course has been offered since 2021 in an online asynchronous mode. All three courses have \href{http://www.sba.oakland.edu/Faculty/isken/teaching.html}{extensive course websites} that are publicly accessible, including access to all of the video content and supporting files.

What kind of Python background do students need to have in order to be ready for the EwP module? A basic familiarity with Python fundamentals for data analysis is required, including:

@@

The EwP module is made up of three submodules. The first of these focuses on building and using a typical spreadsheet model using Python. Then the functionality developed in the first submodule is deployed as a reusable package. The final submodule is an introduction to using Python to manipulate Excel files.

All of the Python modules in the PCDA course as well as the entire AAP course are taught using a combination of Jupyter notebooks, Python scripts, and videos walking through the notebooks. Jupyter notebooks, part of \href{https://jupyter.org/}{Project Jupyter}, are an example of \textit{literate computing} \cite{Jupyter_proposal} and allow one to create educational narratives that support iterative exploration and development. The ability to weave standard explanatory language with formal programming languages has made notebook computing a defacto standard in the data science education community \cite{perkel2018jupyter}. Jupyter notebooks can also include equations written in \LaTeX, data visualizations, and other multi-media. The technology is free, open source, runs on most computing platforms, and has a huge user community that has created a large number of resources for students, faculty and industry practitioners.

Central to Jupyter notebooks is a mix of \textit{code cells} and \textit{markdown cells}. Code cells contain executable code while markdown cells contain plain text using \href{URL}{markdown} for styling text. For example, Figure \ref{cells_edit_model} shows two markdown cells separated by a code cell as they look when in editing mode. After running the cells, they appear as show in Figure \ref{cells_after_running}. Mixing code and markdown cells makes it quite easy to create interactive tutorials in which students can not only run pre-written code, but can also add their own code - all with the aid of supporting explanatory text. All of the blog posts upon which EwP is based were written in Jupyter notebooks. There are countless introductory tutorials for Jupyter notebooks, but the official Project Jupyter Documentations \cite{ProjectJupyterDocumentation} is a very good place to start if you are new to notebook computing.
